### 2021-07-11 docs/manuscript/manuscript.tex
Rephrase text

Despite having only two genes, the regulatory mechanism underlying
the topology 9 generates dynamically more complex expression patterns
compared to other topologies, such as the
I3-FFL; however, and despite the fact that the Gierer-Meinhardt model form
patterns that scale with tissue size~\cite{shoaf1984}, we found that topology 9
was more sensible to changes in the morphogenetic field size than those
topologies based on I3-FFL (\nameref{S2_Table}),
indicating that the additional regulatory interactions need to be fine-tuned
depending on the number of cells in the field.\\

@@

These two events conduce to the repression of the gene C by the gene A in the
rightmost part of the morphogenetic field and by the morphogen in the leftmost
part of the morphogenetic field, thus generating a slightly higher level of
expression of gene C in the center of the field. Because gene C activates its
own expression, a band of expression begins to form. This band of expression
then begins to activate the gene A also in the center of the field
(\nameref{S5_Fig}).
% (Fig~\ref{fig:topol9-dyn}).

@@

This is a complex mechanism remindful of the Gierer-Meinhardt model, but
described by the first time in this study in the context of multiple morphogen
inputs, and it would be interesting to study its feasibility and possible
implementation in synthetic gene regulatory networks.

@@

In this work, we have conducted extensive computational explorations of small
GRNs with the capacity to generate a simple spatial pattern in a 1-D
morphogenetic field. Our findings shed new light on regulatory rules and
dynamical mechanisms that  GRNs observed in nature could potentially
implement to achieve fundamental spatio-temporal tasks, which might
guide for future synthetic biology implementations.

Overall, based on our extensive search space exploration we identified
a great variety of pattern-forming GRN topologies  with distinctive and often
widely shared building blocks (network motifs), which are likely to confer
each topology the ability to achieve specific regulatory tasks in a semi-autonomous
manner. However, we also found a great diversity of network
topologies that could produce a striped gene expression pattern, where 615 out of
714 were multi-input topologies, showing that considering multiple
morphogen inputs allows finding a different solution for the patterning
problem. Most importantly, we found that this ensemble of pattern-forming GRNs
tends to form a highly connected meta-graph, which highlights emergent
properties, such as robustness and evolvability, of complex biological
systems at a very high level~\cite{kitano2004}.\\

Robustness is a defining feature of any developmental patterning process.
Our results suggest that the underlying morphogen interpretation
mechanisms (GRNs) are highly accessible by exploring the
topology space constrained by selection for their final steady-state and
by their ability to buffer noise~\cite{lo_robust_2015, exelby_2021}.
In our study, the most abundant topologies were found forming highly
connected clusters in a so-called neutral network of genotypes,
implying that the corresponding GRN topologies would tend to be robust
to quantitative changes in the strength of regulatory interactions, diffusion, and
degradation parameters, as well as to qualitative changes in the topology
itself. In essence, such organization could be seen as an intrinsic evolutionary
property of GRNs that facilitates the coexistence of robustness and
evolvability~\cite{wagner_robustness_2008}. For instance, bridge topologies
in the neutral network connect distinct classes of pattern-forming
GRNs~(\nameref{S6_Fig}), suggesting that an organism could switch
between different GRNs (for example, between a Gierer-Meinhardt-like network
to an Incoherent Feed-Forward Loop) with minimal fitness costs. This high-level
property of the neutral network of topologies could thus facilitate the fine-tuning
of the regulatory systems through genetic modifications to meet
specific functional requirements under selective pressures.

@@

Other studies have shown that Incoherent Feed-Forward Loops can form
spatial stripes of gene expression~\cite{ishihara_cross_2005}. In particular,
we found that the I3-FFL is one of the network motifs that appears more
frequently in these GRNs, as has also been reported by Cotterell \&
Sharpe~\cite{Cotterell2010}, and has been found in transcriptional networks involved
in cell fate definition~\cite{Li2019}. Moreover, our results indicate that this
is a robust motif that can be useful for constructing synthetic gene
regulatory networks exposed to stochastic fluctuations in the
environment and mutations.

>>>>

Despite having only two genes, the regulatory mechanism underlying
the topology 9 generates dynamically more complex expression patterns
compared to other topologies, such as the
I3-FFL; however, and despite the fact that the Gierer-Meinhardt model forms
patterns that scale with tissue size~\cite{shoaf1984}, we found that topology 9
was more sensible to changes in the morphogenetic field size than those
topologies based on I3-FFL (\nameref{S2_Table}),
indicating that the additional regulatory interactions need to be fine-tuned
depending on the number of cells in the field.\\

@@

These two events lead to the repression of the gene C by the gene A in the
rightmost part of the morphogenetic field and by the morphogen in the leftmost
part of the morphogenetic field, thus generating a slightly higher level of
expression of gene C in the center of the field. Because gene C activates its
own expression, a band of expression begins to form. This band of expression
then begins to activate the gene A also in the center of the field
(\nameref{S5_Fig}).
% (Fig~\ref{fig:topol9-dyn}).

@@

This is a complex mechanism reminiscent of the Gierer-Meinhardt model, which
to the best of our knowledge is reported for the first time in the context of
multiple morphogen inputs, and could thus be of particular interest to for
synthetic biology implementations to assess the feasibility and robustness
conferred by this mechanism to GRNs.

@@

In this work, we have conducted extensive computational explorations of small
GRNs with the capacity to generate a simple spatial pattern in a 1-D
morphogenetic field. Our findings shed new light on regulatory rules and
dynamical mechanisms that GRNs in nature could potentially implement to
achieve stereotyped spatio-temporal tasks.

Overall, based on our extensive search space exploration we identified
a great variety of pattern-forming GRN topologies  with distinctive and often
widely shared building blocks (network motifs), which are likely to confer
each topology the ability to achieve specific regulatory tasks in a semi-autonomous
manner. In particular, we uncovered a great variety of network
topologies that could produce a striped gene expression pattern, where 615 out of
714 were found to be multi-input topologies, which highlights the importance of
often disregarded pleiotropic signaling events as a potentially critical design
specification rule for successful synthetic biology implementations of GRNs
with various types of pattern-forming abilities. Most importantly, we found that
this ensemble of pattern-forming GRNs
tends to form a highly connected meta-graph, which highlights emergent
properties, such as robustness and evolvability, of complex biological
systems at a very high level~\cite{kitano2004}.\\

Robustness is a defining feature of any developmental patterning process.
Our results suggest that the underlying morphogen interpretation
mechanisms (GRNs) are highly accessible by exploring the
topology space constrained by selection for their final steady-state and
by their ability to buffer noise~\cite{lo_robust_2015, exelby_2021}.
In our study, the most abundant topologies were found forming highly
connected clusters in a so-called neutral network of genotypes,
implying that the corresponding GRN topologies would tend to be robust
to quantitative changes in the strength of regulatory interactions, diffusion, and
degradation parameters, as well as to qualitative changes in the topology
itself. In essence, such organization could be seen as an intrinsic evolutionary
property of GRNs that facilitates the coexistence of robustness and
evolvability~\cite{wagner_robustness_2008}. For instance, bridge topologies
in the neutral network connect distinct classes of pattern-forming
GRNs~(\nameref{S6_Fig}), suggesting that evolutionary transitions between
different GRNs (for example, between a Gierer-Meinhardt-like network to an
Incoherent Feed-Forward Loop) could be achieved with minimal fitness costs. This high-level
property of the neutral network of topologies could thus facilitate the fine-tuning
of the regulatory systems through genetic modifications to meet
specific functional requirements under selective pressures.

@@

Other studies have shown that Incoherent Feed-Forward Loops can form
spatial stripes of gene expression~\cite{ishihara_cross_2005}. In particular,
we found that the I3-FFL is one of the network motifs that appears more
frequently in these GRNs, which is in agreement with the study by Cotterell \&
Sharpe~\cite{Cotterell2010}, and has been found in transcriptional networks involved
in cell fate definition~\cite{Li2019}. Moreover, our results indicate that this
is a robust motif that can be useful for constructing synthetic gene
regulatory networks exposed to stochastic fluctuations in the
environment and mutations.
